---
title: "TCGA Citokines dataset"
output: html_notebook
---


## expression association 



```{r}
dataset_ssgsea = "/liulab/asahu/data/ssgsea/xiaoman/TCGA_ALLTPM.txt"

dataset_ssgsea = fread(dataset_ssgsea)
expression_mat= t(as.matrix(dataset_ssgsea[,seq(2,ncol(dataset_ssgsea)),with=F]))
head(rownames(expression_mat))

colnames(expression_mat) = dataset_ssgsea$gene_name
patient.name = colnames(dataset_ssgsea)[-1]
patient.name = gsub(patient.name, pattern="-", replacement=".")
patient.name = substring(patient.name, 1, 16)
rownames(expression_mat) = patient.name
head(patient.name)
head(sample.name$x)
expression_match = expression_mat[sample.name$x,]
expression_match_norm = avinash::normalize.expression(expression_match)

all.genes = colnames(expression_match_norm)
```


```{r}
source("../epoch-142/source.proj.R")
tcga.dataset = fread("~/project/deeplearning/icb/data/tcga/scrna.v4.allgenes.nopcs/dataset.txt")
sample.name = fread("~/project/deeplearning/icb/data/tcga/scrna.v4.allgenes.nopcs/samples_name.txt")
tcga.dataset.match = tcga.dataset[match(rownames(expression_match_norm), sample.name$x),]
citokines = fread("/liulab/xmwang/oxphos_proj/loading_data/surface/cytokine.txt", header=F)
surface.genes = fread("/liulab/xmwang/oxphos_proj/loading_data/surface/ExpressionLigRec.txt", header=T)
citokines = intersect(citokines$V2, all.genes)
surface.genes = setdiff(intersect(surface.genes$ApprovedSymbol, all.genes), citokines)
# intersect(citokines$V2, surface.genes$ApprovedSymbol)
# setdiff(citokines$V2, surface.genes$ApprovedSymbol)
length(citokines)
length(surface.genes)
if.start.inx = which(colnames(tcga.dataset)== "B_cells_naive")
dataset.new = cbind(tcga.dataset.match[,.(cancertype)], expression_match_norm[,c(citokines,surface.genes)], tcga.dataset.match[,seq(if.start.inx, ncol(tcga.dataset.match)), with=F])
output.dir = "~/project/deeplearning/icb/data/tcga/citokines.v1/"
write.dataset(output.dir = output.dir, dataset = dataset.new, sample.name = rownames(expression_match_norm))
file.copy("~/project/deeplearning/icb/data/Getz_scRNA/TCR.AA.V2/params.json", output.dir)
file.copy("~/project/deeplearning/icb/data/Getz_scRNA/TCR.AA.V2/datasets_tsne_list.txt", output.dir)
file.copy("~/project/deeplearning/icb/data/Getz_scRNA/TCR.AA.V2/datasets_test_list.txt", output.dir)

dataset.new.col = colnames(dataset.new)
monitored.mutation = c("ARID1A", "ARID1B", "SMARCA4",  "STK11", "B2M", "PTEN", "PTPN2", "APLNR")
which( dataset.new.col %in%  paste0(monitored.mutation, ".mut")) -2 
length(input.output.genes) + 4
length(input.output.genes) + 4 + dim(if.data)[2] + dim(protien.data)[2]  + length(monitored.mutation)
```

```{r}
dataset_ssgsea = "/liulab/asahu/data/ssgsea/xiaoman/TCGA_ALLTPM.txt"

dataset_ssgsea = fread(dataset_ssgsea)
expression_mat= t(as.matrix(dataset_ssgsea[,seq(2,ncol(dataset_ssgsea)),with=F]))
head(rownames(expression_mat))

colnames(expression_mat) = dataset_ssgsea$gene_name
patient.name = colnames(dataset_ssgsea)[-1]
patient.name = gsub(patient.name, pattern="-", replacement=".")
patient.name = substring(patient.name, 1, 16)
rownames(expression_mat) = patient.name
head(patient.name)
head(sample.name$x)
expression_match = expression_mat[sample.name$x,]
expression_match_norm = avinash::normalize.expression(expression_match)

all.genes = colnames(expression_match_norm)
```



## For Xiaoman
```{r}
oxphos.dataset =  dataset.new[,1346:1403,with=F]
oxphos.dataset$sample.name =  rownames(expression_match_norm)
save(file="~/project/deeplearning/icb/data/tcga/oxphos.dataset.RDATA", oxphos.dataset)

```

## read output and check if there is more correlation in outputs. 

```{r}
deepimmune.out = fread("~/project/deeplearning/icb/data/tcga/citokines.v1/tensorboardLog/20190924-011458/best_val_0.csv")
deepimmune.out.col = colnames(deepimmune.out)
inp.col = deepimmune.out.col[10:136]
dat= deepimmune.out[,inp.col, with=F]

xx = cor(dat, method = "spearman", use = "pairwise.complete.obs"); xx[is.na(xx)] = 0
hc = hclust(as.dist(1 - abs(xx)), method = "complete")
M = xx[order.dendrogram(as.dendrogram(hc)), order.dendrogram(as.dendrogram(hc))]
xx = cor(t(dat), method = "spearman", use = "pairwise.complete.obs"); xx[is.na(xx)] = 0
hr = hclust(as.dist(1 - abs(xx)), method = "complete")

library(corrplot)
corrplot(M, method = "circle", tl.cex=.4)

dat.out= deepimmune.out[,paste0(inp.col,".output"), with=F]
setnames(dat.out, seq_along(inp.col), inp.col)
xx = cor(dat.out, method = "spearman", use = "pairwise.complete.obs"); xx[is.na(xx)] = 0
M1 = xx[order.dendrogram(as.dendrogram(hc)), order.dendrogram(as.dendrogram(hc))]
corrplot(M1, method = "circle", tl.cex=.4)

    library(ComplexHeatmap)
    Heatmap(dat, name = name, column_title = "Variables", row_title = "Samples", 
        row_names_gp = gpar(fontsize = 7))
    
require(heatmap3)
 heatmap3(dat, Rowv = as.dendrogram(hr), Colv = as.dendrogram(hc), 
        scale = "col", balanceColor = T, showRowDendro = F, showColDendro = F, 
        cexRow = 0.5, cexCol = 0.5)

```



## Immune genes
```{r}

immune.genes = avinash::get.immune.genes()   
tcga.immune.genes = expression_match[,immune.genes]

nontranscriptome.immune.faactors = tcga.dataset[,c(102:103, 110:123, 128:137),with=F]                                     
tcga.immune.factors = tcga.dataset[,c(66:137, 178:188),with=F]


```




## Protein association
Check association with at least one immune fator.
```{r}
rrpa = fread("/liulab/asahu/data/ssgsea/xiaoman/TCGA_gdc/TCGA-RPPA-pancan-clean.txt")
rrpa[,SampleID:=gsub(substring(SampleID,1,16), pattern="-", replacement = ".")]
rrpa.match = rrpa[match(sample.name$x, SampleID)]
rrpa.mat = rrpa.match[,-(1:2),with=F]
temp = calc.cor(rrpa.mat, nontranscriptome.immune.factors, use = "pairwise.complete.obs")



temp2 = calc.cor(rrpa.mat, tcga.immune.factors, use = "pairwise.complete.obs")
aa = temp2$estimate
aa = aa[which(apply(aa, 1, sd, na.rm=T) > 0), which(apply(aa, 2, sd, na.rm=T) > 0)] 


temp3 = calc.cor(rrpa.mat, tcga.immune.genes, use = "pairwise.complete.obs")
aa = temp3$estimate
aa = aa[which(apply(aa, 1, sd, na.rm=T) > 0), which(apply(aa, 2, sd, na.rm=T) > 0)] 


protien.sel1 = which(rowSums(abs(temp$estimate)> 0.3 , na.rm = T) > 0)
protien.sel2 = which(rowSums(abs(temp2$estimate)> 0.4 , na.rm = T) > 0)
protien.sel3 = which(rowSums(abs(temp2$estimate)> 0.4 , na.rm = T) > 0)
protien.sel = unique(c(names(protien.sel1), names(protien.sel2), names(protien.sel3)))


rrpa.dataset = rrpa.mat[,protien.sel,with=F]
setnames(rrpa.dataset, seq(ncol(rrpa.dataset)), paste(colnames(rrpa.dataset), "prot", sep=".")) 

cors.list = cor(rrpa.dataset, expression_match, use = "pairwise.complete.obs")
cors.list = cors.list[colnames(rrpa.dataset),]
aaa = data.table(rownames(cors.list), seq(nrow(cors.list)))
topmost.cors.genes = sapply(aaa$V1, function(tt) {
  xx = cors.list[tt,] 
  xx = xx[abs(xx) > 0.15]
  names(which.max(abs(xx)))
}
)

top.cors = lapply(aaa$V1, function(tt) {
  xx = cors.list[tt,] 
  xx = xx[order( abs(xx), decreasing = T)[1:10]] 
  xx[abs(xx) >0.15]
  names(xx)
}
)
rrpa.topmost.cors = unique(unlist(topmost.cors.genes))
rrpa.top.cors = unique(unlist(top.cors))
```

## Mutation data(Havel et al. 2019) : 

STK11, B2M, PTEN, PTPN2, APLNR, SWI/SNF (PBAF)
```{r}
mut.genes = c("STK11", "B2M", "PTEN", "PTPN2", "APLNR")
baf.complex.genes=fread("/liulab/asahu/data/ssgsea/xiaoman/baf_complex.csv")$V2
baf.complex.genes =baf.complex.genes[-1]
all.mut.genes = unique(c(mut.genes, baf.complex.genes))

mutation = fread("/liulab/asahu/data/ssgsea/xiaoman/TCGA_gdc/TCGA_consolidated.abs_mafs_truncated.fixed.txt")
head(mutation$Hugo_Symbol)
# names(table(mutation.dt$Consequence))
synonymous.mut.type = c("downstream_gene_variant",   "intron_variant",       "stop_retained_variant",    "synonymous_variant",       "upstream_gene_variant")   

mutation[,SampleID:=gsub(substring(sample,1,16), pattern="-", replacement = ".")]
mut.samples = unique(mutation$SampleID)
mutation.dt = mutation[Hugo_Symbol %in% all.mut.genes]
mutation.dt = mutation.dt[!(Consequence %in% synonymous.mut.type)]
mutation.dt =mutation.dt[SampleID%in%sample.name$x]
mut.bin.mat = table(mutation.dt[,.( SampleID , Hugo_Symbol)])
mut.ccf.mat = xtabs(ccf_CI95_low~ SampleID + Hugo_Symbol, data=mutation.dt)
mut.cff.mat = mut.ccf.mat[,colnames(mut.bin.mat)]

mut.bin = matrix(NA, nrow=length(sample.name$x), ncol = ncol(mut.bin.mat), dimnames = list(sample.name$x, colnames(mut.bin.mat)))
mut.bin[intersect(mut.samples, sample.name$x),] =0
mut.bin[rownames(mut.bin.mat),] = mut.bin.mat

mut.ccf = matrix(NA, nrow=length(sample.name$x), ncol = ncol(mut.ccf.mat), dimnames = list(sample.name$x, colnames(mut.ccf.mat)))
mut.ccf[intersect(mut.samples, sample.name$x),] =0
mut.ccf[rownames(mut.ccf.mat),] = mut.ccf.mat
colnames(mut.ccf) = paste0(colnames(mut.ccf.mat), ".ccf")
mut.dataset = cbind( mut.ccf, mut.bin)
genes.curr1 = names(sort(colSums(mut.bin.mat),decreasing = T))[1:6]
genes.curr = unique(c(mut.genes, genes.curr1))

cors.list = cor(mut.dataset, expression_match, use = "pairwise.complete.obs")
rownames(cors.list) = colnames(mut.dataset)
aaa = data.table(rownames(cors.list), seq(nrow(cors.list)))
aaa = aaa[V1 %in% c(genes.curr, paste0(genes.curr, ".ccf"))]


topmost.cors.genes = sapply(aaa$V1, function(tt) {
  xx = cors.list[tt,] 
  xx = xx[abs(xx) > 0.1]
  names(which.max(abs(xx)))
}
)

top.cors = lapply(aaa$V1, function(tt) {
  xx = cors.list[tt,] 
  xx = xx[order( abs(xx), decreasing = T)[1:10]] 
  xx[abs(xx) >0.1]
  names(xx)
}
)
mut.topmost.cors = unique(unlist(topmost.cors.genes))
mut.top.cors = unique(unlist(top.cors))

```

##aggregation of genes
PCA
```{r}
topmost.cors.genes  = unique(c(if.topmost.cors, rrpa.topmost.cors, mut.topmost.cors))
top.cors.genes  = unique(c(if.top.cors, rrpa.top.cors, mut.top.cors))
top.cors.genes = setdiff(top.cors.genes, topmost.cors.genes)

top.cors.genes = setdiff(top.cors.genes, colnames(tcga.dataset))

tcga.exp.norm = normalize.expression(expression_match, num.samp.thr =0) 
mat = apply(tcga.exp.norm[,top.cors.genes], 2, function(tt) (tt-mean(tt))/sd(tt))
pcs=tcga.dataset[,2:51,with=F]
residual.pca = function(mat, control){
  apply(mat, 2, function(tt) {
    control$cov = tt
    (lm(cov~., data= control))$residuals
  })
}

uu = residual.pca(mat, control = pcs)
top.gene.pcs = prcomp(uu)
save(file="~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/top.gene.pcs.RData", top.gene.pcs)
ww = (top.gene.pcs$sdev^2)
ww1 = cumsum(ww/sum(ww))
ww1[1:100] # only 30% of explained variance is captured
## save referecence expression 
# ref.expression = expression_match
# ref.cancertype = tcga.dataset$cancertype

save(file="~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/ref.expression.RData", ref.expression, ref.cancertype)
save(file="~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/ref.expression.norm.RData",  ref.expression.norm, ref.cancertype)
# xx = princomp(mat)
# ww = (xx$sdev^2)
# ww1 = cumsum(ww/sum(ww))

```
## Create dataset and write dataset files 
```{r}

top.pcs = top.gene.pcs$x[,1:50]
tcga.cols = colnames(tcga.dataset)
colnames(top.pcs) = paste0("PC", 51:100)
pc.data = cbind(tcga.dataset[,1:51,with=F], top.pcs)

extra.genes.inx = c("TGFB1", "TGFBR2", "KLRC1")
top.genes.extra = c("PMS2", "MSH6", "EPCAM", "MSH2")
input.output.genes = unique(c(extra.genes.inx, top.genes.extra, tcga.cols[c(62, 89:90, 189:510)])) #101
input.only.genes = setdiff(topmost.cors.genes, input.output.genes)
if.cols = tcga.cols[c(66:88, 91:137, 178:188)]
if.cols = setdiff(if.cols, input.output.genes)
if.data  = tcga.dataset[,if.cols, with=F] #78
setnames(if.data, 68:78, paste0("ERV", 1:11))

protien.data = rrpa.dataset
mut.data = mut.dataset
colnames(mut.data) = paste0(colnames(mut.data), ".mut")
survival.data = tcga.dataset[,c(170:177), with=F]

dataset.new = cbind(pc.data, tcga.exp.norm[,input.only.genes], tcga.exp.norm[,input.output.genes], if.data, protien.data, mut.data, survival.data)
output.dir = "~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/"
write.dataset(output.dir = output.dir, dataset = dataset.new, sample.name = sample.name)

```
### Genes instead of top pcs 

```{r}
tcga.cols = colnames(tcga.dataset)
save(file="~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/meta.info.RData", tcga.cols, topmost.cors.genes, top.cors.genes)
```

```{r}
pc.data = tcga.dataset[,1:51,with=F]

extra.genes.inx = c("TGFB1", "TGFBR2", "KLRC1")
top.genes.extra = c("PMS2", "MSH6", "EPCAM", "MSH2")
input.output.genes = unique(c(extra.genes.inx, top.genes.extra, tcga.cols[c(62, 89:90, 189:510)])) #101
input.only.genes = setdiff(unique(c(topmost.cors.genes, top.cors.genes)), input.output.genes)
if.cols = tcga.cols[c(66:88, 91:137, 178:188)]
if.cols = setdiff(if.cols, input.output.genes)
if.data  = tcga.dataset[,if.cols, with=F] #78
setnames(if.data, 68:78, paste0("ERV", 1:11))

protien.data = rrpa.dataset
mut.data = mut.dataset
colnames(mut.data) = paste0(colnames(mut.data), ".mut")
survival.data = tcga.dataset[,c(170:177), with=F]

dataset.new = cbind(pc.data, tcga.exp.norm[,input.only.genes], tcga.exp.norm[,input.output.genes], if.data, protien.data, mut.data, survival.data)
dataset.new.col = colnames(dataset.new)
which(dataset.new.col == input.output.genes[1])
monitored.mutation = c("ARID1A", "ARID1B", "SMARCA4",  "STK11", "B2M", "PTEN", "PTPN2", "APLNR")
which( dataset.new.col %in%  paste0(monitored.mutation, ".mut")) -2 
length(input.output.genes) + 4
length(input.output.genes) + 4 + dim(if.data)[2] + dim(protien.data)[2]  + length(monitored.mutation)
output.dir = "~/project/deeplearning/icb/data/tcga/scrna.v4.genes/"
write.dataset(output.dir = output.dir, dataset = dataset.new, sample.name = sample.name)
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/params.json", "~/project/deeplearning/icb/data/tcga/scrna.v4.genes/.")
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/datasets_tsne_list.txt", "~/project/deeplearning/icb/data/tcga/scrna.v4.genes/.")

```

## All genes no pcs 
Is PC is helping? Are large number of outputs are sufficient to learn large number of parameters? 

```{r}
pc.data = tcga.dataset[,1,with=F]  ## this is cancertype only
tcga.cols = colnames(tcga.dataset)
extra.genes.inx = c("TGFB1", "TGFBR2", "KLRC1")
top.genes.extra = c("PMS2", "MSH6", "EPCAM", "MSH2")
input.output.genes = unique(c(extra.genes.inx, top.genes.extra, tcga.cols[c(62, 89:90, 189:510)])) #101
input.only.genes = setdiff(all.genes, input.output.genes)
if.cols = tcga.cols[c(66:88, 91:137, 178:188)]
if.cols = setdiff(if.cols, input.output.genes)
if.data  = tcga.dataset[,if.cols, with=F] #78
setnames(if.data, 68:78, paste0("ERV", 1:11))

protien.data = rrpa.dataset
mut.data = mut.dataset
colnames(mut.data) = paste0(colnames(mut.data), ".mut")
survival.data = tcga.dataset[,c(170:177), with=F]

dataset.new = cbind(pc.data, tcga.exp.norm[,input.only.genes], tcga.exp.norm[,input.output.genes], if.data, protien.data, mut.data, survival.data)
dataset.new.col = colnames(dataset.new)
temp.dt = data.table(dataset.new.col, seq_along(dataset.new.col))
which(dataset.new.col == input.output.genes[1])
monitored.mutation = c("ARID1A", "ARID1B", "SMARCA4",  "STK11", "B2M", "PTEN", "PTPN2", "APLNR")
which( dataset.new.col %in%  paste0(monitored.mutation, ".mut")) -2 
length(input.output.genes) + 4
length(input.output.genes) + 4 + dim(if.data)[2] + dim(protien.data)[2]  + length(monitored.mutation)
output.dir = "~/project/deeplearning/icb/data/tcga/scrna.v4.allgenes.nopcs/"
write.dataset(output.dir = output.dir, dataset = dataset.new, sample.name = sample.name)
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.genes/params.json", "~/project/deeplearning/icb/data/tcga/scrna.v4.allgenes.nopcs/")
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.genes/datasets_tsne_list.txt", "~/project/deeplearning/icb/data/tcga/scrna.v4.allgenes.nopcs/.")


```
##  Nir's data processing
* PC
* pc.sel
* gene.expression

### Create function of automatically caclulate PC

```{r}

#' Create function of automatically caclulate PC
#'
#' @param dataset_ssgsea_sel Expression matrix
#' @param ref.expression.RData 
#' @param ref.expression.RData 
#' @param cancer.type if any cancer type. Default (NULL) uses all reference.
#' @param len_selected Number of PCs 
#' @param pca_obj.RData Default is NULL so it will reestiamte the PCS. 
#'
#' @return
#' @export
#'
#' @examples
get.pc1 = function(dataset_ssgsea_sel, ref.expression.RData, ref.cancertype.RData = NULL, cancer.type=NULL, len_selected =50, pca_obj.RData = NULL){
  require(avinash)
  if(is.null(pca_obj.RData)){
    pca.obj = NULL
  }else{
    load(pca_obj.RData)
  }
  dataset_ssgsea_norm = normalize.expression(dataset_ssgsea_sel)
  load(ref.expression.RData)
  if(!is.null(ref.cancertype.RData)) load(ref.cancertype.RData)
  
  dataset_ssgsea_sel = dataset_ssgsea_sel[,colnames(ref.expression)]
  stopifnot(identical(colnames(dataset_ssgsea_sel), colnames(ref.expression)))
  if(!is.null(cancer.type)) ref.expression = ref.expression[ref.cancertype ==cancer.type, ]
  dataset_ssgsea_sel = match.expression.distribution(dataset_ssgsea_sel, ref.expression)
  temp_out = get_pca(dataset_ssgsea_sel, pca_obj = pca_obj, scale=F) 
  pca_obj = temp_out$pca_obj
  pca_obj$len_selected = len_selected
  temp_out$pca_out[,seq(pca_obj$len_selected)]
}

```


```{r}

#' Title
#'
#' @param common.genes common genes to imputed 
#' @param incomplete_exp_mat incomplete expression matrix sample x genes
#' @param ref  reference expression matrix sample x genes
#'
#' @return
#' @export
#'
#' @examples
impute.closest.gene = function(common.genes, incomplete_exp_mat, ref= NULL, subsample=NULL){
	  genes.imputed = setdiff(common.genes, colnames(incomplete_exp_mat))
	  if(length(genes.imputed) > 0) {
	  gene1 = colnames(incomplete_exp_mat)
	  if(is.null(class(ref))){
	  	ref.dt = fread("/liulab/asahu/data/ssgsea/xiaoman/Genetech_expression_TPM.txt")
	  	ref = t(as.matrix(ref.dt[,seq(2,ncol(ref.dt)),with=F]))
	  	setnames(ref.dt, 1, "gene_name")
	  	colnames(ref ) = ref.dt$gene_name
	  }
	  ref.dt.gene_name = colnames(ref)
	  if (!is.null(subsample)) {
	      ref = ref[sample.int(nrow(ref), size=nrow(ref)*subsample),]
	  }
	  impute = ref[,genes.imputed, drop=F]
	  only.genes = intersect(gene1, ref.dt.gene_name)
	  incomplete_exp_mat = incomplete_exp_mat[,only.genes, drop=F]
	  exp.present = ref[,only.genes,drop=F]
	  cors = cor(impute, exp.present, use="pairwise.complete.obs")
	  # cors = pcor(impute, exp.present, use="pairwise.complete.obs")
	  genes.inx = apply(cors,1, 
	  	function(tt) ifelse(sum(!is.na(tt)), which.max(tt), NA)
	  	)

	  imputed = incomplete_exp_mat[,genes.inx, drop=F]
	  imputed[is.na(imputed)] = 0
	  colnames(imputed) = genes.imputed
	  merged = cbind(incomplete_exp_mat, imputed) 
	  incomplete_exp_mat = merged
	}
	incomplete_exp_mat[,common.genes]
}

```

```{r}
recalculate.pcs = F
dataset_ssgsea =  "/liulab/asahu/data/ssgsea/xiaoman/getz/dca/mean_norm.tsv"
dataset_phenotype =  "~/project/deeplearning/icb/data/Getz_scRNA/data/cell_label.csv"
dataset_phenotype = "/liulab/asahu/data/ssgsea/xiaoman/getz/cell_label.csv"
dataset_ssgsea_temp = fread(dataset_ssgsea)
dataset_phenotype = fread(dataset_phenotype)
dataset_ssgsea_mat= t(as.matrix(dataset_ssgsea_temp[,seq(2,ncol(dataset_ssgsea_temp)),with=F]))
setnames(dataset_ssgsea_temp, 1, "gene_name")
colnames(dataset_ssgsea_mat) = dataset_ssgsea_temp$gene_name
rownames(dataset_ssgsea_mat)[16292]
dataset_ssgsea_mat = dataset_ssgsea_mat[-16292,] # last column is NAs
rownames(dataset_ssgsea_mat) = colnames(dataset_ssgsea_temp)[-1]
common.genes = all.genes


dataset_ssgsea_mat1 = impute.closest.gene(common.genes,dataset_ssgsea_mat, ref =expression_match, subsample = .1) ## this is from avinash library
dataset_ssgsea_mat = dataset_ssgsea_mat1[ ,common.genes] 
stopifnot(any(!is.na(dataset_ssgsea_mat)))

patient.name = rownames(dataset_ssgsea_mat)
patient.name = gsub(patient.name, pattern="-", replacement=".")
rownames(dataset_ssgsea_mat) = patient.name
setnames(dataset_phenotype, 1, "patient.name")
dataset_phenotype$patient.name = gsub(dataset_phenotype$patient.name, pattern="-", replacement=".")
only_in_phenotype = setdiff(dataset_phenotype$patient.name, patient.name)
only_in_ssgsea = setdiff( patient.name, dataset_phenotype$patient.name)
curr.common.patient = intersect(patient.name, dataset_phenotype$patient.name)
dataset_ssgsea_sel = dataset_ssgsea_mat[match(curr.common.patient, patient.name), ] 
phenotype_sel = dataset_phenotype[match(curr.common.patient, dataset_phenotype$patient.name)]

colnames(phenotype_sel) = gsub(colnames(phenotype_sel), pattern=" ", replacement="_")
colnames(phenotype_sel) = gsub(colnames(phenotype_sel), pattern="-", replacement="_")
phenotype_sel.mod = phenotype_sel

if(recalculate.pcs){
general.pcs = get.pc1(dataset_ssgsea_sel, 
ref.expression.RData = "~/project/deeplearning/icb/data/tcga/scrna.v1/ref.expression.RData",
ref.cancertype.RData = "~/project/deeplearning/icb/data/tcga/scrna.v1/ref.cancertype.RData",
pca_obj.RData = "~/project/deeplearning/icb/data/tcga/scrna.v1//pca_obj.RData"
                       )
}else{
  general.pcs = fread("~/project/deeplearning/icb/data/Getz_scRNA/dca/dataset.txt")[,2:51,with=F]
}

```



##  Expression inputs
```{r}

dataset_ssgsea_sel.norm = normalize.expression(dataset_ssgsea_sel, num.samp.thr = 0) 
# save(file="~/project/deeplearning/icb/results/sessions/temp.RData", dataset_ssgsea_sel.norm, tcga.exp.norm)
# xx = local({load("~/project/deeplearning/icb/results/sessions/temp1.RData"); environment()})
# dataset_ssgsea_sel.norm = xx$dataset_ssgsea_sel.norm
# rm(xx)
dataset_ssgsea_sel.norm = match.expression.distribution(dataset_ssgsea_sel.norm, tcga.exp.norm)

pc.data = data.table(cbind(cancertype="SKCM", general.pcs))
extra.genes.inx = c("TGFB1", "TGFBR2", "KLRC1")
top.genes.extra = c("PMS2", "MSH6", "EPCAM", "MSH2")
input.output.genes = unique(c(extra.genes.inx, top.genes.extra, tcga.cols[c(62, 89:90, 189:510)])) #101
input.only.genes = setdiff(unique(c(topmost.cors.genes, top.cors.genes)), input.output.genes)

nir.sample.name = rownames(dataset_ssgsea_sel)
nir.dataset.new = cbind(pc.data, dataset_ssgsea_sel.norm[,input.only.genes], dataset_ssgsea_sel.norm[,input.output.genes])
output.dir="~/project/deeplearning/icb/data/Getz_scRNA/scrna.v4.genes"
write.dataset(output.dir = output.dir, dataset = nir.dataset.new, sample.name = nir.sample.name)
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.genes/params.json", output.dir)
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/datasets_tsne_list.txt", output.dir)

# save.image("~/project/deeplearning/icb/results/sessions/process.dataset.imputation.tcga.v2.today.RData")
```

## All expression inputs 
```{r}

extra.genes.inx = c("TGFB1", "TGFBR2", "KLRC1")
top.genes.extra = c("PMS2", "MSH6", "EPCAM", "MSH2")
input.output.genes = unique(c(extra.genes.inx, top.genes.extra, tcga.cols[c(62, 89:90, 189:510)])) #101
input.only.genes = setdiff(all.genes, input.output.genes)

nir.sample.name = rownames(dataset_ssgsea_sel)
nir.dataset.new1 = cbind(dataset_ssgsea_sel.norm[,input.only.genes], dataset_ssgsea_sel.norm[,input.output.genes])
nir.dataset.new = data.table(cbind(cancertype="SKCM", nir.dataset.new1))
output.dir="~/project/deeplearning/icb/data/Getz_scRNA/scrna.v4.allgenes"
write.dataset(output.dir = output.dir, dataset = nir.dataset.new, sample.name = nir.sample.name)
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.genes/params.json", output.dir)
file.copy("~/project/deeplearning/icb/data/tcga/scrna.v4.pcs/datasets_tsne_list.txt", output.dir)
rm(nir.dataset.new)
rm(nir.dataset.new1)
save.image("~/project/deeplearning/icb/results/sessions/process.dataset.imputation.tcga.v2.today.v2.RData")
```

